```{r 10_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# Local Regression & GAMs

## Learning Goals {-}

- Clearly describe the local regression algorithm for making a prediction
- Explain how bandwidth (span) relate to the bias-variance tradeoff
- Describe some different formulations for a GAM (how the arbitrary functions are represented)
- Explain how to make a prediction from a GAM
- Interpret the output from a GAM

<br>

Slides from today are available [here](https://docs.google.com/presentation/d/1WOz5gkBDqob_tRzNwJWBrn5wNyfI8n78SGVRACEWgVE/edit?usp=sharing).


<br><br><br>


## GAMs - Options for Fitting {-}

### GAMs (splines + OLS) {-}

We've already talked about how to fit GAM models with splines (`step_ns()`) using the `lm` engine (ordinary least squares).

### GAMs (LOESS) {-}

The `gam` package provides tools for building GAMs with local regression (LOESS). We won't explore this option further (via code) in class because there isn't a `tidymodels` interface.

### GAMs (smoothing splines) in `tidymodels` {-}

Today, we'll try fitting GAM models with smoothing splines in `tidymodels`.

To build GAMs (using smoothing splines) in `tidymodels`, first load the package:

```{r}
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
tidymodels_prefer() # Resolves conflicts, prefers tidymodel functions
```

Then adapt the following code:

```{r}
gam_spec <- 
    gen_additive_mod() %>%
    set_args(select_features = TRUE, adjust_deg_free = tune()) %>%
    set_engine(engine = "mgcv") %>%
    set_mode("regression")

# The implementation of GAMs in tidymodels is different than other methods...
# ...we don't specify a recipe
gam_wf <- workflow() %>% 
    add_variables(outcomes = YOUR_OUTCOME, predictors = c(PREDICTOR1, PREDICTOR2)) %>% 
    add_model(gam_spec, formula = YOUR_OUTCOME ~ CATEGORICAL_PREDICTORS + s(QUANTITATIVE_PREDICTOR, k=10))
# s(x1, k = 10): This makes the smoothing spline have 10 knots

tuning_param_grid <- grid_regular(
    adjust_deg_free(range = c(0.25, 4)),
    levels = 10
)

data_cv <- vfold_cv(___, v = 10)

tune_output <- tune_grid( 
    gam_wf,
    resamples = data_cv,
    metrics = metric_set(mae),
    grid = tuning_param_grid
)
```

<br>

**Picking the best tuning parameter and visualizing the GAM**

```{r}
# Select best model & fit to full training data
best_param <- tune_output %>% select_best()
best_by_1se_param <- tune_output %>% 
    select_by_one_std_err(metric = "mae", desc(adjust_deg_free))

gam_mod_best <- finalize_workflow(gam_wf, best_param) %>%
    fit(data = ___)

gam_mod_best1se <- finalize_workflow(gam_wf, best_by_1se_param) %>%
    fit(data = ___)

# Plot functions for each predictor
# Dashed lines are +/- 2 SEs
fit_gam_model %>% pluck("fit") %>% plot() 
```



<br><br><br>



## Exercises {-}

**You can download a template RMarkdown file to start from [here](template_rmds/10-loess-gams.Rmd).**

Before proceeding, install the `mgcv` package by entering `install.packages("mgcv")` in the Console.

We'll continue using the `College` dataset in the `ISLR2` package to explore models for graduation rate. You can use `?College` in the Console to look at the data codebook.

```{r}
library(ISLR2)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
tidymodels_prefer() # Resolves conflicts, prefers tidymodel functions

data(College)

# A little data cleaning
college_clean <- College %>% 
    mutate(school = rownames(College)) %>% 
    filter(Grad.Rate <= 100) # Remove one school with grad rate of 118%
rownames(college_clean) <- NULL # Remove school names as row names
```


### Exercise 1: Conceptual warmup {-}

a. How does high/low span relate to bias and variance of a local regression (LOESS) model?

b. How does high/low lambda relate to bias and variance of a smoothing spline in a GAM model?

c. Do you think that a GAM with all possible predictors will have better or worse performance than an ordinary (fully linear) least squares model with all possible predictors? Explain your thoughts.

d. Why might we want to perform variable selection before fitting a GAM? How could stepwise selection or LASSO help with this?





### Exercise 2: Local regression (LOESS) {-}

Use LOESS (`geom_smooth()`) to explore the relationship between `Apps` and `Grad.Rate` for different values of `span` (between 0 and 1).

- How is varying span like varying number of neighbors in KNN?
- How would you describe the relationship between number of applications and graduation rate?

```{r}
college_clean %>%
    ggplot(aes(x = Apps, y = Grad.Rate)) + 
    geom_point(alpha = 0.2) + # Adjust point transparency with alpha
    geom_smooth(span = 0.4, method = "loess", se = FALSE) + # vary span
    xlim(c(0,20000)) +
    theme_classic()
```





### Exercise 3: Building a GAM in `tidymodels` {-}

Suppose that initial variable selection investigations have given us a set of predictors to include in our GAM. The code below for building a GAM is complete (nothing to fill in), but before looking at output, we are going to closely examine this code in comparison to the code for our previous methods: LASSO and KNN.

- The `adjust_deg_free` argument is like lambda in LASSO. Higher values of this tuning parameter mean that "wiggliness" is penalized more.
- The `select_features = TRUE` allows for the ability to eliminate a predictor via penalization (more likely with higher `adjust_deg_free`).

```{r}
set.seed(123)
gam_spec <- 
    gen_additive_mod() %>%
    set_args(select_features = TRUE, adjust_deg_free = tune()) %>%
    set_engine(engine = "mgcv") %>%
    set_mode("regression")

# The implementation of GAMs in tidymodels is different than other methods...
# ...we don't specify a recipe
gam_wf <- workflow() %>% 
    add_variables(outcomes = Grad.Rate, predictors = c(Private, Apps, Top10perc, P.Undergrad, Outstate, perc.alumni)) %>% 
    add_model(gam_spec, formula = Grad.Rate ~ Private + s(Apps, k=10) + s(Top10perc, k=10) + s(P.Undergrad, k=10) + s(Outstate, k=10) + s(perc.alumni, k=10))

tuning_param_grid <- grid_regular(
    adjust_deg_free(range = c(0.25, 4)),
    levels = 10
)

data_cv8 <- vfold_cv(college_clean, v = 8)

# This takes a few seconds
tune_output <- tune_grid( 
    gam_wf,
    resamples = data_cv8,
    metrics = metric_set(mae),
    grid = tuning_param_grid
)
```

We can take a look at the test error metrics from CV and choose an optimal tuning parameter:

```{r}
tune_output %>% collect_metrics()
tune_output %>% show_best()

autoplot(tune_output) + theme_classic()

# Select best model & fit to full training data
best_param <- tune_output %>% select_best()
best_by_1se_param <- tune_output %>% 
    select_by_one_std_err(metric = "mae", desc(adjust_deg_free))

gam_mod_best <- finalize_workflow(gam_wf, best_param) %>%
    fit(data = college_clean)

gam_mod_best1se <- finalize_workflow(gam_wf, best_by_1se_param) %>%
    fit(data = college_clean)
```

Let's visualize the estimated nonlinear functions.

- What about these plots indicates that using a GAM instead of ordinary linear regression was probably a good choice?
- Compare the plots resulting from the two different choices for "best" tuning parameter. Which choice would you prefer and why?
- For each of the plots, write a sentence describing what you learn about the relationship between graduation rate and that predictor.

```{r}
# Plot the estimated nonlinear functions for all predictors...
# ...for the GAM resulting from the adjust_deg_free parameter value that gave the lowest error
gam_mod_best$fit$fit$fit %>% plot(all.terms = TRUE, pages = 1)

# ...for the GAM resulting from the adjust_deg_free parameter value that gave the lowest error
gam_mod_best1se$fit$fit$fit %>% plot(all.terms = TRUE, pages = 1)
```


