```{r 21_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# Clustering (Project Work)

## Learning Goals {-}

- Implement k-means and hierarchical clustering for your project dataset and interpret findings

<br>

Slides from today are available [here](https://docs.google.com/presentation/d/1Iy9ZmNaS-XpGO2B0An-2eG2YP-e9hmIdfDOrx_ZWrQI/edit?usp=sharing).


<br><br><br>


## Dataset choice {-}

If not working on your project dataset, feel free to choose one of the following three datasets to work with:

1. Wine Attributes (download [here](https://bcheggeseth.github.io/253_fall_2021/template_rmds/wine.csv))
    - 178 Italian wines were analyzed
    - Variables (from Chemical Analysis)
        - Alcohol
        - Malic acid
        - Ash
        - Alcalinity of ash
        - Magnesium
        - Total phenols
        - Flavanoids
        - Nonflavanoid phenols
        - Proanthocyanins
        - Color intensity
        - Hue
        - OD280/OD315 of diluted wines
        - Proline

```{r}
library(readr)
wine <- read_csv("wine.csv")
```

2. Mall Customers (download [here](https://bcheggeseth.github.io/253_fall_2021/template_rmds/mall_customers.csv))
    - 200 individuals
    - Variables
        - Binary Gender
        - Age
        - Annual Income (in $1000's)
        - Spending Score (summary of buying behavior)
  
```{r}
library(readr)
customers <- read_csv("mall_customers.csv")
```  
  
3. Credit Card Clients (download [here](https://bcheggeseth.github.io/253_fall_2021/template_rmds/creditcard.csv))
    - Almost 9000 credit card holders
    - Variables based on 6 months of time
        - CUSTID: Identification of Credit Card holder
        - BALANCE : Balance amount left in their account to make purchases
        - BALANCEFREQUENCY : How frequently the Balance is updated, score between 0 and 1 (1 = frequently updated, 0 = not frequently updated)
        - PURCHASES : Number of purchases made from account
        - ONEOFFPURCHASES : Maximum purchase amount done in one-go
        - INSTALLMENTSPURCHASES : Amount of purchase done in installment
        - CASHADVANCE : Cash in advance given by the user
        - PURCHASESFREQUENCY : How frequently the Purchases are being made, score between 0 and 1 (1 = frequently purchased, 0 = not frequently purchased)
        - ONEOFFPURCHASESFREQUENCY : How frequently Purchases are happening in one-go (1 = frequently purchased, 0 = not frequently purchased)
        - PURCHASESINSTALLMENTSFREQUENCY : How frequently purchases in installments are being done (1 = frequently done, 0 = not frequently done)
        - CASHADVANCEFREQUENCY : How frequently the cash in advance being paid
        - CASHADVANCETRX : Number of Transactions made with "Cash in Advanced"
        - PURCHASESTRX : Number of purchase transactions made
        - CREDITLIMIT : Limit of Credit Card for user
        - PAYMENTS : Amount of Payment done by user
        - MINIMUM_PAYMENTS : Minimum amount of payments made by user
        - PRCFULLPAYMENT : Percent of full payment paid by user
        - TENURE : Tenure of credit card service for user

```{r}
library(readr)
credit <- read_csv("creditcard.csv")
```

### Analysis {-}

Whether you are using your project dataset or one of the above datasets, the goals and analysis plans are as follows:

- Goal: Cluster the data to gain insights into latent groupings and patterns

- Available clustering methods
    - K-means with all quantitative variables
    - Partitioning around medoids (pam) as a robust version of K-means
        - If you have at least one categorical variable, `daisy()` will calculate [Gower's distance](https://stat.ethz.ch/R-manual/R-devel/library/cluster/html/daisy.html). (Gower's distance automatically handles scaling.)
    - Hierarchical clustering
        - If you have at least one categorical variable, use `daisy()` to calculate Gower's distance.

When using `daisy()`, you will need to make sure that all categorical variables are of the `factor` type in R:

```{r}
data <- data %>%
    mutate(
        cat_var1 = as.factor(cat_var1),
        cat_var2 = as.factor(cat_var2)
    )
```

```{r}
# K-means with all quantitative variables
kmeans(data, centers = k)
```

```{r}
# PAM for a robust version of k-means when there are some categorical variables
# You will need to install the "cluster" package to use daisy()
library(cluster)
pam(daisy(data), k = k)

# Choosing an appropriate number of clusters
# Create storage vector for total within-cluster sum of squares
tot_wc_ss <- rep(0, 15) 

# Loop
for (k in 1:15) {
    # Perform clustering
    pam_out <- pam(daisy(mtcars), k = k)

    # Store the total within-cluster sum of squares
    tot_wc_ss[k-1] <- sum(pam_out$clusinfo[,"av_diss"]*pam_out$clusinfo[,"size"])
}

plot(1:15, tot_wc_ss, xlab = "Number of clusters", ylab = "Total within-cluster sum of squares")
```

```{r}
# Hierarchical clustering on...
# ...all quantitative variables
hclust(dist(scale(data)), method = "CHOOSE_LINKAGE_TYPE")

# ...a mix of quantitative and categorical variables
hclust(daisy(data), method = "CHOOSE_LINKAGE_TYPE")
```

- General process
    1. Decide which variables you want to use in your clustering and why. Use `select(your_data, chosen_var1, chosen_var2, etc)` to choose this subset of variables.
    2. Implement k-means (if all variables are quantitative) or PAM (if some variables are categorical).
        - Decide on a reasonable number of clusters by inspecting a plot like in Exercise 6 of Topic 19.
        - Running these methods multiple times (to see if the results are consistent despite the random initialization steps) might not be possible if computational time is high for your dataset.
    3. Implement hierarchical clustering.
        - Visualizing the dendrogram might be tough if you have a lot of cases. You may need to rely on just cutting the tree and inspecting the characteristics of the resulting clusters.
        - Use the same number of clusters as from k-means/PAM as a starting point for the number of clusters here.
    4. Insights
        - Use summary statistics like in the Topic 19 exercises and visualizations like in the Topic 20 exercises to understand the features of the clusters.


<br><br>


## Clustering in the Wild {-}

To give you a taste of how these methods get use in "the wild" world of science, here are a few papers (quality varies):

- Image Segmentation (https://link.springer.com/article/10.1007/s11042-021-10594-9)
- Bacteria Clustering (https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0002843)
- Document Clustering (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7790388/)
- Clustering ICD10 Diagnosis Codes (https://arxiv.org/abs/1909.00306)
- Clustering Activity Sequences (https://www.sciencedirect.com/science/article/abs/pii/S0968090X21000395?via%3Dihub)



<br><br>



## R Coding challenges {-}

The best way to learn new things about R is to work on a data project.

- The goals drive what code is needed. 
- Learn them as you need them. 


What things have come up so far for you?

- What has been the most frustrating?
- When do you get stuck?
- What are you wanting to do with your data?


Besides class projects, you can practice visualizing data:

- [TidyTuesday Challenges](https://github.com/rfordatascience/tidytuesday)
    - Check out David Robinson's [TidyTuesday's Screencasts](https://www.youtube.com/watch?v=nx5yhXAQLxw)
  


